---
title: "p8105_hw2_jx2562"
author: "Jingyi"
output: github_document
---

```{r, message=F}
library(tidyverse)
```

## Problem 1

In this problem, we aim to merge three data sets, _pols-month.csv_, _unemployment.csv_, and _snp.csv_ into one. 

Clean the _pols-month.csv_.
```{r clean_pols-month, message=F}
pm_df =
  read_csv("data/pols-month.csv") |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    month = as.numeric(month),
    year = as.numeric(year))|>
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "party") |>
  filter(party != 0) |>
  select(-c(party, day)) |>
  arrange(year,month) |>
  mutate(month = month.abb[month]) |>
  janitor::clean_names()

tail(pm_df)

```

Clean the _snp.csv_ with the similar process as above. Mainly focus on the representation of date.
```{r clean_snp, message=F}
snp_df =
  read_csv("data/snp.csv") |>
  separate(date, into = c("month", "day", "year"), sep = "/") |>
  mutate(
    month = as.numeric(month),
    year = as.numeric(year),
    year = ifelse(year > 25, 1900 + year, 2000 + year)) |>
  select(year, month, close) |>
  arrange(year,month) |>
  mutate(month = month.abb[month]) |>
  janitor::clean_names()

tail(snp_df)

```

Clean the _unemployment.csv_ so that it could fit in the previous two datasets.

```{r clean_unemployment, message=F}
uep_df =
  read_csv("data/unemployment.csv") |>
  pivot_longer(
    cols = Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) |>
  janitor::clean_names()

tail(uep_df)
```

Here the three data frames are all lead by year and month with a correct order. Next, we merge the three data frames together.

```{r merge_data, message=F}
merge_df = pm_df |>
  full_join(snp_df) |>
  full_join(uep_df)

tail(merge_df)

```


