---
title: "p8105_hw2_jx2562"
author: "Jingyi"
output: github_document
---

```{r, message=F}
library(tidyverse)
library(readxl)
```

## Problem 1

In this problem, we aim to merge three data sets, _pols-month.csv_, _unemployment.csv_, and _snp.csv_ into one. 

Clean the _pols-month.csv_.
```{r clean_pols-month, message=F}
pm_df =
  read_csv("data/pols-month.csv") |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    month = as.numeric(month),
    year = as.numeric(year))|>
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "party") |>
  filter(party != 0) |>
  select(-c(party, day)) |>
  arrange(year,month) |>
  mutate(month = month.abb[month]) |>
  janitor::clean_names()

head(pm_df)

```

Clean the _snp.csv_ with the similar process as above. Mainly focus on the representation of date.
```{r clean_snp, message=F}
snp_df =
  read_csv("data/snp.csv") |>
  separate(date, into = c("month", "day", "year"), sep = "/") |>
  mutate(
    month = as.numeric(month),
    year = as.numeric(year),
    year = ifelse(year > 25, 1900 + year, 2000 + year)) |>
  select(year, month, close) |>
  arrange(year,month) |>
  mutate(month = month.abb[month]) |>
  janitor::clean_names()

head(snp_df)

```

Clean the _unemployment.csv_ so that it could fit in the previous two datasets.

```{r clean_unemployment, message=F}
uep_df =
  read_csv("data/unemployment.csv") |>
  pivot_longer(
    cols = Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) |>
  janitor::clean_names()

head(uep_df)
```

Here the three data frames are all lead by year and month with a correct order. Next, we merge the three data frames together.

```{r merge_data, message=F}
crisis_df = pm_df |>
  left_join(snp_df) |>
  left_join(uep_df)

head(crisis_df)

```

The **crisis_df** is combined by the three data frames, **pm_df**, **snp_df**, and **uep_df**, by `left_join()`. In this way, all values in the three data frames are kept. The merged data frame has _`r ncol(crisis_df)`_ columns and _`r nrow(crisis_df)`_ rows. It has the variables **`r colnames(crisis_df)`**. The recorded time of the data ranges from _`r paste0(crisis_df$year[1],sep = "-", crisis_df$month[1])`_ to _`r paste0(crisis_df$year[nrow(crisis_df)], sep = "-", crisis_df$month[nrow(crisis_df)])`_.

The **pm_df** has _`r ncol(pm_df)`_ columns and _`r nrow(pm_df)`_ rows related to the number of national politicians who are democratic or republican at any given time. It contains variables **`r colnames(pm_df)`**. And the time ranges from _`r paste0(pm_df$year[1],sep = "-", pm_df$month[1])`_ to _`r paste0(pm_df$year[nrow(pm_df)], sep = "-", pm_df$month[nrow(pm_df)])`_.

The **snp_df** has _`r ncol(snp_df)`_ columns and _`r nrow(snp_df)`_ rows related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. It contains variables **`r colnames(snp_df)`**. And the time ranges from _`r paste0(snp_df$year[1],sep = "-", snp_df$month[1])`_ to _`r paste0(snp_df$year[nrow(snp_df)], sep = "-", snp_df$month[nrow(snp_df)])`_.

The **uep_df** has _`r ncol(uep_df)`_ columns and _`r nrow(uep_df)`_ rows related the unemployment rate at a given time. It contains variables **`r colnames(uep_df)`**. And the time ranges from _`r paste0(uep_df$year[1],sep = "-", uep_df$month[1])`_ to _`r paste0(uep_df$year[nrow(uep_df)], sep = "-", uep_df$month[nrow(uep_df)])`_.


## Problem 2

First, define a path to the dataset. 

```{r}
path_to_data = "./data/202309 Trash Wheel Collection Data.xlsx"
```

Read and clean the Mr. Trashwheel dataset. 

```{r}
mr_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) |>
	janitor::clean_names() |> 
	drop_na(dumpster) |>
  mutate(
    homes_powered = round(as.numeric(weight_tons)*500/30, 2),
    type = "Mr.",
    year = as.numeric(year)) |>
  select(-dumpster) 

head(mr_df)
```

Read and clean the Professor Trashwheel dataset. 

```{r}
prof_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Professor Trash Wheel",
		range = cell_cols("A:M")) |>
	janitor::clean_names() |> 
	drop_na(dumpster) |>
  mutate(
    homes_powered = round(as.numeric(weight_tons)*500/30, 2),
    type = "Professor",
    year = as.numeric(year)) |>
  select(-dumpster)

head(prof_df)
```

Read and clean the Gwynnda Trashwheel dataset. 

```{r}
gwynnda_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Gwynnda Trash Wheel",
		range = cell_cols("A:K")) |>
	janitor::clean_names() |> 
	drop_na(dumpster) |>
  mutate(
    homes_powered = round(as.numeric(weight_tons)*500/30, 2),
    type = "Gwynnda",
    year = as.numeric(year)) |>
  select(-dumpster)

head(prof_df)
```

Merge the three data sets.

```{r}
trashwheel_df = mr_df |>
  bind_rows(prof_df) |>
  bind_rows(gwynnda_df)

trashwheel_df =
  select(trashwheel_df, month:homes_powered, plastic_bags, type) |>
  arrange(year, month)

head(trashwheel_df)
```

This **trashwheel_df** merged information from the Mr., Professor and Gwynnda Trashwheel trash collector in Baltimore, Maryland. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of _`r nrow(trashwheel_df)`_ rows in our final dataset. Additional data sheets include month precipitation data.

In these datasets:

* The variables in the final data frame **trashwheel_df** are **`r colnames(trashwheel_df)`**.

*  The duration of the data collected in the final data set is from _`r min(trashwheel_df$date)`_ to _`r max(trashwheel_df$date)`_. By checking the _year_ variable of the starting date, there are some errors in data collection. Despite the outlier, the corrected time duration is from _`r pmin(trashwheel_df$date)[2]`_ to _`r max(trashwheel_df$date)`_.

*  The mean weight in tons of trash collected during these years is `r trashwheel_df |> pull(weight_tons) |> sum() |> round(2)`.

*  The total weight of trash collected by Professor Trash Wheel is `r prof_df |> pull(weight_tons) |> sum()`.

*  The total number of cigarette butts collected by Gwynnda Trash Wheel in July of 2021 is `r gwynnda_df |> filter(year == 2021 & month == "July") |> pull(cigarette_butts) |> sum()/1000` thousands.


## Problem 3

First, we read and clean the baseline data.

```{r}
baseline_df = 
  read_csv("data/MCI_baseline.csv", skip = 1,show_col_types = FALSE) |>
  janitor::clean_names()

baseline_filter = baseline_df |>
  mutate(
    sex = ifelse(sex == 1, "male", "female"),
    apoe4 = ifelse(apoe4 == 1, "carrier", "non-carrier"),
  ) |>
  filter(age_at_onset=="." | current_age < age_at_onset)

head(baseline_filter)
```

The samples are filtered by MCI development. Those participants who did not have MCI by the date of latest record are discarded. 
The data has features are stated below:

*  In the beginning, a total of `r nrow(baseline_df)` participants was recruited, with `r baseline_df |> filter(sex == 1) |> nrow()` males and `r baseline_df |> filter(sex == 0) |> nrow()` females. Despite the participants who developed MCI before the study, we have `r nrow(baseline_filter)`

*  Among these participants, `r baseline_filter |> filter(age_at_onset != ".") |> nrow()` participants developed MCI by the end of follow-up, with  `r baseline_filter |> filter(sex == "male") |> nrow()` males and `r baseline_filter |>  filter(sex == "female") |> nrow()` females.

*  The average age of all data collected is `r round(mean(baseline_df$current_age), 2)`. The average age of filtered samples when the study began is `r round(mean(baseline_filter$current_age), 2)`. The average MCI onset age in the filtered sample is `r baseline_filter |> filter(age_at_onset != ".") |> pull(age_at_onset) |> as.numeric() |> mean() |> round(2)` 

*  The proportion of women as APOE4 carriers is `r round(nrow(baseline_filter[baseline_filter$sex == "female" & baseline_filter$apoe4 == "carrier",]) / nrow(baseline_filter[baseline_filter$sex == "female",]), 2)` 


Clean the biomarker data

```{r, message = F}
biomarker_df = 
  read_csv("data/mci_amyloid.csv", skip = 1, show_col_types = FALSE) |>
  janitor::clean_names() |>
  select(id = study_id, everything()) |> 
  mutate(
    baseline = as.numeric(baseline),
    time_8 = as.numeric(time_8)
  )

head(biomarker_df)
```

The longitudinally observed biomarker values are loaded. The name of the column `study_id` was changed to `id` for easier data sets merging in the later steps.

*  The longitudinal amyloid β 42/40 ratio of `r nrow(biomarker_df)` participants were collected.

*  The variables include `r colnames(biomarker_df)`, with `r ncol(biomarker_df) - 1` time points.

*  The average ratio at the baseline is `r round(mean(biomarker_df$baseline, na.rm = T), 4)`. The average ratio at the time 8 is `r round(mean(biomarker_df$time_8, na.rm = T), 4)`.


Check the different participants in the two sets.

```{r}
only_baseline = setdiff(baseline_filter$id, biomarker_df$id)
only_biomarker = setdiff(biomarker_df$id, baseline_filter$id)

```

There are `r length(only_baseline)` participants only appears in the baseline dataset. They are # `r only_baseline` participants. `r length(only_biomarker)` participants only appears in the amyloid dataset. They are # `r only_biomarker` participants.


Merge the demographic information and biomarker information for participants in both data sets.

```{r, message = F}
MCI_df = 
  inner_join(baseline_filter, biomarker_df)

head(MCI_df)
write.csv(MCI_df, "data/MCI_summary.csv")
```

The `MCI_df` combines the information from `baseline_filter` and `biomarker_df`. The details of the data frame is stated below:

*   It has `r nrow(MCI_df)` observations and `r ncol(MCI_df)` variables.

*   Their average age at the beginning of the study is `r round(mean(MCI_df$current_age), 2)`, ranging from `r min(MCI_df$current_age)` to `r max(MCI_df$current_age)` .

*   Their average age at the onset of MCI is `r MCI_df |> filter(age_at_onset != ".") |> pull(age_at_onset) |> as.numeric() |> mean() |> round(2)`, ranging from `r MCI_df |> filter(age_at_onset != ".") |> pull(age_at_onset) |> as.numeric() |> min()` to `r max(MCI_df$age_at_onset)` .

*   Among a total of `r nrow(MCI_df)` participants, `r MCI_df |> filter(sex == "male") |> nrow()` are males and `r MCI_df |> filter(sex == "female") |> nrow()` are females.

*   `r MCI_df |> filter(apoe4 == "carrier") |> nrow()` are APOE4 carriers and `r MCI_df |> filter(apoe4 == "non-carrier") |> nrow()` are non-carriers.

*   The amyloid β 42/40 ratio at the baseline has a average of `r round(mean(biomarker_df$baseline, na.rm = T), 4)`, ranging from `r min(MCI_df$baseline, na.rm=T)` to `r max(MCI_df$baseline, na.rm=T)` .

*   The amyloid β 42/40 ratio at the time 8 has a average of `r round(mean(biomarker_df$time_8, na.rm = T), 4)`, ranging from `r min(MCI_df$time_8, na.rm=T)` to `r max(MCI_df$time_8, na.rm=T)` .




