---
title: "p8105_hw2_jx2562"
author: "Jingyi"
output: github_document
---

```{r, message=F}
library(tidyverse)
library(readxl)
```

## Problem 1

In this problem, we aim to merge three data sets, _pols-month.csv_, _unemployment.csv_, and _snp.csv_ into one. 

Clean the _pols-month.csv_.
```{r clean_pols-month, message=F}
pm_df =
  read_csv("data/pols-month.csv") |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    month = as.numeric(month),
    year = as.numeric(year))|>
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "party") |>
  filter(party != 0) |>
  select(-c(party, day)) |>
  arrange(year,month) |>
  mutate(month = month.abb[month]) |>
  janitor::clean_names()

head(pm_df)

```

Clean the _snp.csv_ with the similar process as above. Mainly focus on the representation of date.
```{r clean_snp, message=F}
snp_df =
  read_csv("data/snp.csv") |>
  separate(date, into = c("month", "day", "year"), sep = "/") |>
  mutate(
    month = as.numeric(month),
    year = as.numeric(year),
    year = ifelse(year > 25, 1900 + year, 2000 + year)) |>
  select(year, month, close) |>
  arrange(year,month) |>
  mutate(month = month.abb[month]) |>
  janitor::clean_names()

head(snp_df)

```

Clean the _unemployment.csv_ so that it could fit in the previous two datasets.

```{r clean_unemployment, message=F}
uep_df =
  read_csv("data/unemployment.csv") |>
  pivot_longer(
    cols = Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) |>
  janitor::clean_names()

head(uep_df)
```

Here the three data frames are all lead by year and month with a correct order. Next, we merge the three data frames together.

```{r merge_data, message=F}
crisis_df = pm_df |>
  full_join(snp_df) |>
  full_join(uep_df)

head(crisis_df)

```

The **crisis_df** is combined by the three data frames, **pm_df**, **snp_df**, and **uep_df**, by full_join(). In this way, all values in the three data frames are kept. The merged data frame has _`r ncol(crisis_df)`_ columns and _`r nrow(crisis_df)`_ rows. It has the variables **`r colnames(crisis_df)`**. The recorded time of the data ranges from _`r paste0(crisis_df$year[1],sep = "-", crisis_df$month[1])`_ to _`r paste0(crisis_df$year[nrow(crisis_df)], sep = "-", crisis_df$month[nrow(crisis_df)])`_.

The **pm_df** has _`r ncol(pm_df)`_ columns and _`r nrow(pm_df)`_ rows related to the number of national politicians who are democratic or republican at any given time. It contains variables **`r colnames(pm_df)`**. And the time ranges from _`r paste0(pm_df$year[1],sep = "-", pm_df$month[1])`_ to _`r paste0(pm_df$year[nrow(pm_df)], sep = "-", pm_df$month[nrow(pm_df)])`_.

The **snp_df** has _`r ncol(snp_df)`_ columns and _`r nrow(snp_df)`_ rows related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. It contains variables **`r colnames(snp_df)`**. And the time ranges from _`r paste0(snp_df$year[1],sep = "-", snp_df$month[1])`_ to _`r paste0(snp_df$year[nrow(snp_df)], sep = "-", snp_df$month[nrow(snp_df)])`_.

The **uep_df** has _`r ncol(uep_df)`_ columns and _`r nrow(uep_df)`_ rows related the unemployment rate at a given time. It contains variables **`r colnames(uep_df)`**. And the time ranges from _`r paste0(uep_df$year[1],sep = "-", uep_df$month[1])`_ to _`r paste0(uep_df$year[nrow(uep_df)], sep = "-", uep_df$month[nrow(uep_df)])`_.


## Problem 2

First, define a path to the dataset. 

```{r}
path_to_data = "./data/202207 Trash Wheel Collection Data.xlsx"
```

Read and clean the Mr. Trashwheel dataset. 

```{r}
mr_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) |>
	janitor::clean_names() |> 
	drop_na(dumpster) |>
  mutate(
    homes_powered = round(as.numeric(weight_tons)*500/30, 2),
    type = "Mr.",
    year = as.numeric(year)) |>
  select(-dumpster) 

head(mr_df)
```

Read and clean the Professor Trashwheel dataset. 

```{r}
prof_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Professor Trash Wheel",
		range = cell_cols("A:M")) |>
	janitor::clean_names() |> 
	drop_na(dumpster) |>
  mutate(
    homes_powered = round(as.numeric(weight_tons)*500/30, 2),
    type = "Professor",
    year = as.numeric(year)) |>
  select(-dumpster)

head(prof_df)
```

Read and clean the Gwynnda Trashwheel dataset. 

```{r}
gwynnda_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Gwynnda Trash Wheel",
		range = cell_cols("A:K")) |>
	janitor::clean_names() |> 
	drop_na(dumpster) |>
  mutate(
    homes_powered = round(as.numeric(weight_tons)*500/30, 2),
    type = "Gwynnda",
    year = as.numeric(year)) |>
  select(-dumpster)

head(prof_df)
```

Merge the three data sets

```{r}
trashwheel_df = mr_df |>
  bind_rows(prof_df) |>
  bind_rows(gwynnda_df)

trashwheel_df =
  select(trashwheel_df, month:homes_powered, plastic_bags, type) |>
  arrange(date)

head(trashwheel_df)
```

